---
title: "Midterm II Solution"
author: "Yidong Zhou"
date: "5/21/2021"
output: pdf_document
---

## Q1

```{r}
plastic <- read.table("http://users.stat.ufl.edu/~rrandles/sta4210/Rclassnotes/data/textdatasets/KutnerData/Chapter%20%201%20Data%20Sets/CH01PR22.txt")
colnames(plastic) <- c('Y', 'X')
str(plastic)
n <- nrow(plastic)
p <- 2
```

### (a) 15 points

- $(\mathbf{X}'\mathbf{X})^{-1}$

Here we have only one predictor. The design matrix $\mathbf{X}$ should be $n$ by $2$, where $n$ is the number of observations ($n=16$ in this case).

```{r}
X <- cbind(rep(1, n), plastic$X)# n by 2
Y <- plastic$Y# n by 1
solve(t(X)%*%X)
```

- $\mathbf{b}$

To make sure that `b` is a vector, you can either use `as.vector()` or `as.numeric()`. See `class(solve(t(X)%*%X)%*%t(X)%*%Y)`.

```{r}
b <- as.vector(solve(t(X)%*%X)%*%t(X)%*%Y)
b
```

- SSE

The same for SSE, `as.vector()` or `as.numeric()` can be used.

```{r}
H <- X%*%solve(t(X)%*%X)%*%t(X) # n by n
I <- diag(n)
SSE <- as.vector(t(Y)%*%(I-H)%*%Y)
MSE <- SSE/(n-p)
SSE
```

### (b) 10 points

$s\{b_i\}$ is the square root of $s^2\{b_i\}$, i.e., the square root of the $(i, i)$ entry of the variance covariance matrix of $\mathbf{b}$. $s\{b_i, b_j\}$ is the $(i, j)$ entry of the variance covariance matrix of $\mathbf{b}$.

```{r}
s2b <- MSE*solve(t(X)%*%X)
sqrt(diag(s2b)[1])
sqrt(diag(s2b)[2])
s2b[1, 2]# or s2b[2, 1]
```

### (c) 5 points

Note that $\frac{1}{n}\mathbf{J}$ is equal to $\mathbf{1}(\mathbf{1}'\mathbf{1})^{-1}\mathbf{1}'$, where $\mathbf{1}$ is the $n$-vector of ones.

```{r}
J <- rep(1, n)%*%t(rep(1, n))
H-J/n
```

### (d) 10 points

See $(6.50)$ in the textbook.

```{r}
alpha <- 1-0.95
c(L = b[2] - qt(1-alpha/2, n-p)*sqrt(diag(s2b)[2]), 
  U = b[2] + qt(1-alpha/2, n-p)*sqrt(diag(s2b)[2]))
```


## Q2

## Code Appendix
```{r, ref.label=knitr::all_labels(), echo=TRUE, eval=FALSE}
```
